Seasonality in governing upper-ocean dynamics at submesoscales

### 2016-06-07 writeup/rocha_etal.tex
Updates abstract abd intro

\keypoints{\item Upper-ocean submesoscale (10-100 km) turbulence and innertia-gravity waves
                 undergo strong seasonal cycles that are out-of-phase.
           \item Submesoscale turbulence dominates the horizontal  velocity and
                  sea-surface height variability in late Winter/early Spring.
            \item Inertia-gravity waves are summertime enhanced near the surface and
                  dominate the surface variability in late Summer/early Fall.
                  }

@@

\begin{abstract}
    We show that internal waves are a loud noise.
\end{abstract}

@@

Recent interest in upper-ocean dynamics has focused on the strong seasonal
cycle of shallow baroclinic instabilities and their role in submesoscale (1-100 km)
turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
have also suggested that inertia-gravity waves contribute significantly
to the near-surface variability at submesoscales \citep{richman_etal2012,
buhler_etal2014,rocha_etal2016}.

Using the output of two new primitive equation $1/24^\circ$ and $1/24^\circ$ global
 simulations with embedded tides, we show that the near-surface
expression of inertia-gravity waves undergoes a strong seasonal cycle that is out-of-phase
with the seasonal cycle of the balanced variability. This seasonal cycle is associated
with changes in the upper-ocean
stratification. Consistent with previous studies,
deep late Winter mixed-layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
The horizontally divergent part of the flow, however, peaks in late Summer/early Fall,
when the upper-ocean is strongly stratified. Hence these results suggest a
strong seasonal modulation of the dominant upper-ocean submesoscale dynamics:
quasi-balanced turbulence in late Winter/early Spring and
quasi-linear inertia-gravity waves in late Summer/early Fall.

@@

@@

 is significant: the RMS is $\delta/f > 0.1$).
Most of these horizontal divergent flows are supra-inertial and project
on scales smaller than 100 km: daily-averaging or smoothing the velocity fields
 (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
Interestingly, the seasonal cycle in horizontal divergence is out-of-phase with the
seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
Summer/early Fall than it is in late Winter/early Spring. The divergence associated
with submesoscale subinertial eddies is small and its seasonal cycle is in phase
with that of vorticity (see green lines).  The seasonal cycle of the horizontal
divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
stratification.
\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

>>>>

\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
                  undergo strong seasonal cycles that are out-of-phase.
           \item  Submesoscale turbulence dominates the horizontal velocity and
                  sea-surface height variability in late Winter/early Spring.
            \item Inertia-gravity waves dominate the the horizontal velocity and
                   sea-surface height variability in late Summer/early Fall.
                  }

@@

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km). Consistent with recent studies, deep late Winter mixed layers
    are prone to baroclinic instabilities and submesoscale turbulence
    prevails in wintertime. Inertia-gravity waves, on the other hand, are enhanced near
    the surface when the upper ocean in strongly stratified. Hence, in summertime,
    inertia-gravity waves strongly dominate the surface kinetic energy and
    sea-surface height variance at submesoscales.
\end{abstract}

@@

Recent interest in upper-ocean dynamics has focused on the strong seasonal
cycle of shallow baroclinic instabilities and their role in submesoscale roughly 1-100 km)
turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
have also suggested that inertia-gravity waves contribute significantly
to the near-surface variability at submesoscales \citep{richman_etal2012,
buhler_etal2014,rocha_etal2016}, but their seasonality has not been investigated.

Using the output of two new $1/24^\circ$ and $1/48^\circ$ global
numerical simulations with embedded tides, we show that inertia-gravity waves undergo
a strong seasonality near the surface in the Kuroshio Extension region.
The seasonal cycle of inertia-gravity waves is out-of-phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
deep late Winter mixed layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.

Inertia-gravity waves, on the other hand, peak in late Summer/early Fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late Winter/early Spring,
whereas quasi-linear inertia-gravity waves prevail in late Summer/early Fall.
Because submesoscale turbulence is significantly diminished in late Summer/early Fall,
the present results suggest that inertia-gravity waves account for most of the
summertime submesoscale SSH variability.

@@

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

 is significant: the RMS is $\delta/f > 0.1$).
Most of these horizontal divergent flows are supra-inertial and project
on scales smaller than 100 km: daily-averaging or smoothing the velocity fields
 (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
Interestingly, the seasonal cycle in horizontal divergence is out-of-phase with the
seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
Summer/early Fall than it is in late Winter/early Spring. The divergence associated
with submesoscale subinertial eddies is small and its seasonal cycle is in phase
with that of vorticity (see green lines).  The seasonal cycle of the horizontal
divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
stratification.
